### 2021-08-21 appendix/appendix.Rmd
appendix structure

---
title: |
       | Supplementary Materials for "Electoral Systems, Competition, and Incentives for Corruption"
author:
  - Umberto Mignozzetti[^umberto]
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 11pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
always_allow_html: true
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>.

\appendix

```{r, include=FALSE}
# Knitr options
knitr::opts_chunk$set(fig.pos = "H") # holds figure position
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r, include=FALSE}
## Starting
#XXXX

# Required packages
#XXXX

# Load datasets
#XXXX
```

\newpage

# Formal Model
\label{app:specs}

Consider three sets of players: a set of incumbents $ I \ = \ \{ I_1, \ \cdots, \ I_K \}$, a set of challengers $C \ = \ \{ C_1, \ \cdots, \ C_L \}$, and a continuum of $V$ voters.

Each incumbent has a budget of $A$ units, that has to be invested in public goods. The level of expenditure necessary to provide the goods is a random variable $\eta \in [0,A]$, where $\eta \sim f(\eta)$, and $f$ is a weakly increasing PDF function with CDF $F$. If the public goods are successfully provided, we say that the policy result is good ($\psi = G$). Otherwise, we say the policy result is bad ($\psi = B$).

Let $\e$ be the vector of expenditures enacted by the incumbents. The public goods are provided whenever $K \cdot \eta \leq \sum_i e_i$, where $K$ represents number of incumbents. I assume that all incumbents observe the true $\eta$, and it is common-knowledge among them.

After choosing the expenditures, the public goods provision is realized and is common knowledge to all players. Now the campaign stage starts. Challengers face no running costs, and although they do not observe $\eta$, they know $\e$ and $\psi$.\footnote{This assumption does not change the game because an entry cost would only shift the equilibrium, preserving its qualitative characteristics.}

Upon observing $\e$ and $\psi$, challengers update their beliefs about $\eta$, and decide simultaneously weather to use negative campaign against the incumbents, paying a cost of $c \in \mathbb{R}_{+}$ whenever they use their full resources to expose a corrupt incumbent. After their decisions, the result is a vector $\dd \in \{0,1\}^{K}$ of sucessfully exposed corrupt incumbents, for each incumbent $i$ in $\{1, \cdots, K\}$. The probability that challenger $j$ investigating $i$ credibly communicate the $i$'s true corruption engagement to the voter is $\lambda: [0,A] \rightarrow [0,1]$. I assume $\lambda(0) = 0$ and $\lambda^{'} > 0$. Therefore, the probability of being credibly exposed depends on the distance between the optimal expenditure and the amount spend by the incumbent.

Finally, the voters have to decide whether to retain or dismiss the incumbents, upon observing $\psi$ and $\dd$. Their choices are summarized in a vector $\bups \in \{0,1\}^{K+L}$, where $1$ means that the politician is elected (or reelected, in case of an incumbent) and $0$, meaning not elected. We have that $\sum_i \upsilon_i = K$, and when the voter is asked to decide, she cannot change the results of $\psi$ and $\dd$. The voters' decisions are always sequentially rational. Therefore, I impose that they behave optimally by dismissing an incumbent $j$ that were credibly related with corruption, or all incumbents, when $\psi=B$.

The time-line of the game is as follows:

\begin{enumerate}
 \item The nature draws $\eta$;
 \item Each incumbent observes $\eta$ and chooses, simultaneously, $\e$;
 \item $\psi$ is realized. Challengers choose, also simultaneously, whether to invest resources to expose corrupt incumbents;
 \item The vector of successfully exposed incumbents $\dd$ are realized. The voter chooses $\bups(\psi, \ \dd)$ and the game ends.
\end{enumerate}

An incumbent $i$'s strategy is defined as:

\[
\varepsilon_i:[0,A] \rightarrow [0,A] 
\]

And the challenger $j$'s strategy is defined as:

\[
\bvphi_j:\{G, \ B \} \times [0,A]^K \rightarrow [0,1]^K
\]

Representing the amount of time invested by each challenger $j \in \{1, \cdots , L \}$ to expose an incumbent $i$. Notice that $\bvphi_j$ is a vector of size $K$, where each component is the probability of exposing one of the incumbents. For the representative voter we have:

\[
\bups:\{G, \ B \} \times \{0, \ 1 \}^{K} \rightarrow \{0, \ 1 \}^{K+L}
\]

That represents the decision to retain or dismiss each politician, with their replacement. The expected utilities for each player are, for the incumbents:

\[
U_i (\varepsilon_i; \ \eta, \ \bvphi^{*}, \ \bups^{*}, \ \varepsilon_{-i}^{*}) \ = \ u(\varepsilon_i - \eta) + B \ \Prob (\upsilon_i = 1 \ | \ \psi, \dd) 
\]

Where $B$ is the reward for staying in office (assumed the same for every player) and $\Prob (\upsilon_i = 1 \ | \ \psi, \dd)$ is the probability of being retained, given $\psi$ and the exposure profile $\dd$. For the challenger $j$:

\[
U_j (\bvphi_j; \ \bvphi^{*}_{-j}, \ \bups^{*}, \ \bveps^{*}) \ = \ B \int_{0}^A \Prob (\upsilon_j = 1 \ | \ \eta, \bveps^{*}, \bvphi^{*})\mu(\eta|\bveps^{*}) d\mu(\eta) - c \sum_{i=1}^K \varphi_{ij}(\psi, \bveps^{*})
\]

With $c > 0$ being the cost of performing one full corruption exposure. Notice that the total amount spent on exposing corrupt politicians will be the sum of the amounts spent on each exposure. The representative voter's expected utility is defined as:

\[
U_V (\psi, \dd, \bups) \ = \ \indf \{\psi=G \}r_G
\]

Where $\indf$ is the indicator function for $\psi=G$ and $r_G$ the value of the public good for the voter. Notice that no $\psi=B$ is observed on the equilibrium path, as underspending is a dominated strategy.

Also, any retention rule selected for the voters is sequentially rational, because there is no uncertainty about the results when they vote. I choose the optimal rule for the voter as dismiss all incumbents when $\psi=B$ or replace incumbents that were successfully exposed by challengers when $\psi=G$. This rule is ex-post efficient as over-dismissal would make incumbents to pocket all the money regardless of $\eta$.

Finally, I solve for the \textit{Perfect Bayesian Equilibrium} \citep{Fudenberg1991book, Fudenberg1991}. A perfect Bayesian equilibrium consists in a pair of strategies and beliefs such that:

\begin{enumerate}
 \item In each information set, the decision should be sequentially rational;
 \item Each strategy is compatible with its beliefs concerning the behavior on and off the equilibrium path.
\end{enumerate}

Also, as this is a signaling game, I use \textit{universal divinity} as the equilibrium selection criteria \citep{BanksSobel1987}.\footnote{The restriction imposed by the universal divinity requires that the challengers, facing a deviation from the equilibrium path, put probability one on the type that are more likely to deviate. Moreover, as we have a continuum of types in one of the types dimensions, we use the approach by \citet{Malaith1987}, which consists in find the monotonicity in the payoffs and single crossing conditions, characterizing the behavior of the types inside the intervals of the type space, analyzing the incentives to pool or separate among the types.}

\begin{definition}
 \label{def_eqn}
 A Perfect Bayesian Equilibrium for the game described, with $|I|=K$ and $|C|=L$, $K$ and $L$ in $\mathbb{N}$, type space $T=[0,A]$, consists in optimal strategies $\bveps^{*}$, $\bvphi^{*}$, $\bups^{*}$ and beliefs $\mu^*$, that satisfy
 \begin{enumerate}
  \item $\forall t_i \in T$, $\varepsilon^{*}_i(t_i)$ maximizes $ U_i (\varepsilon_i; \ t_i, \ \bvphi^{*}, \ \bups^{*}, \ \varepsilon_{-i}^{*}) $ $\forall i \in \{1, \cdots, K\}$.
  \item $\forall \bveps \in \left[0,A \right]^K $ and $\psi \in \{B, \ G \}$, $\bvphi^{*}_j(\bveps, \psi)$ maximizes $ U_j (\bvphi_j; \ \bvphi^{*}_{-j}, \ \bups^{*}, \ \bveps^{*}, \ \mu^*) $, $\forall j \in \{1, \cdots, L\}$
  \item $\forall \psi \in \{B, \ G \}$ and $\dd \in \{0,1\}^L$, $\bups^{*}(\psi, \dd)$ maximizes $ U_V (\psi, \dd, \bups^*) $.
  \item The beliefs are derived using Bayes rule whenever possible, and for all $\varepsilon_i$ that are off-the-equilibrium path, $\mu^*(t_i| \varepsilon_i)>0$ (in fact, 1) only if $\widehat{\varepsilon}_i \in \arg \min_{t_i} p(t_i, \widehat{\varepsilon}_i)$, where $ p(t_i, \widehat{\varepsilon}_i)$ is the profit from deviating from $\varepsilon_i$ to $\tilde{\varepsilon}_i$ for a type $t_i$.
 \end{enumerate}
\end{definition}

## Single-Member Districts with One Challenger

Let $|I|=1$ and $|C|=1$. Solving the game backwards, the voters retain the incumbent if the policy result was good and no corruption was exposed.

From the challenger's perspective, as there is no cost of entry, she decides whether to invest her campaign resources in exposing the incumbent. She does so when the expected utility of exposing is greater than or equal the expected utility of not exposing, given her beliefs about the true value of $\eta$. Thus:

\[
B \int_{t_i \in T} \Prob (\bups = (0,1)\ | \ \eta, \varepsilon^*, \varphi)\mu^*(t_i|\varepsilon^*) d\mu - c \varphi(\psi, \varepsilon^*) \geq 0
\]

Rearranging and applying the fact that $ \Prob(\bups = (0,1) | \varepsilon^*, \varphi ( \psi, \varepsilon^*), \eta) = \varphi(\psi, \varepsilon^*) \lambda(\varepsilon^*-\eta) $ and $\mu^*(t_i | \varepsilon^*) = \mu(\eta) = \dfrac{f(\eta)}{\int_0^{A}f(\eta)d\eta}$,

\[
B \varphi(\psi, \varepsilon^*) \int_0^A \lambda(\varepsilon^* - \eta) \dfrac{f(\eta)}{\int_0^{A}f(\eta)d\eta} d\eta - c \varphi(\psi, \varepsilon^*)
\]

Isolating $\varphi(\psi, \varepsilon^*)$, letting $\lambda_\varepsilon = \int_0^A \lambda(\varepsilon^* - \eta) \dfrac{f(\eta)}{\int_0^{A}f(\eta)d\eta} d\eta$ and $Q(B, c) = \dfrac{c}{B}$, we see that sequential rationality implies that the challenger expose corrupt incumbents for sure whenever $\lambda_\varepsilon > Q(B, c)$, never expose corruption when $ \lambda_\varepsilon < Q(B, c)$, and is indifferent when $\lambda_\varepsilon = Q(B, c)$.

Now let $\dfrac{\int_0^A \lambda(A - \eta)f(\eta) d\eta}{\int_0^Af(\eta)d\eta} = Q^*$. If $Q \geq Q^*$, we have only one equilibrium, which will be to play $A$ regardless of $\eta$. And there will be no exposures in equilibrium, as they are too expensive to be performed. To avoid this situation, we impose the following condition:

\begin{assumption}\label{assExist}
We assume $B$ and $c$ are restricted in such a way that $Q<Q^*$.
\end{assumption}

Given this assumption, there exists $\overline{\eta}$ such that

\[
\lambda_A = \dfrac{\int_{\overline{\eta}}^A \lambda(A - \eta)f(\eta) d\eta}{\int_{\overline{\eta}}^Af(\eta)d\eta} = Q(B, c)
\]

In equilibrium, we seek for roots to the equation $\lambda_A-Q(B, c)$. These roots are well characterized, as if $\overline{\eta}=0$, $\lambda_A - Q (B, c) \geq 0$ and if $\overline{\eta}=0$, $\lambda_A=0$ and so the value is negative. As this function is continuous, by the intermediate value theorem, we have a root, that we call $\overline{\eta}$. The monotonicity of $f$ also ensures that each root is unique.

$\overline{\eta}$ has an intuitive interpretation: it is the minimal value for $\eta$ such that above this value, the incumbent always prefers to play $A$ regardless of $\eta$.

Now, let a probability function $\delta(\varepsilon)$ to characterize the equilibrium amount of negative campaign when $\varepsilon \in [ 0, \ \overline{\eta}]$. The optimal $\varepsilon$ solves the equation $\lambda_\varepsilon = Q(B, c)$, making the challenger indifferent between exposing or not the incumbent's corruption. As $\lambda(.)$ has a well defined inverse function, the optimal strategy for the incumbent is:

\begin{equation*}
\varepsilon^*(\eta) \ = \ \eta + \lambda^{-1}\left(\dfrac{c}{B}\right)
\end{equation*}

Therefore, the required value plus a kickback $\kappa$ defined as $\kappa = \lambda^{-1}\left(\dfrac{c}{B}\right)$, whenever $\eta \in [0, \overline{\eta}]$, and $A$ otherwise.

Now, reasoning backwards, the incumbent only plays this strategy when it is incentive compatible for her, and to be an optimal response, it should be an $\arg \max$ of $U_I$. Taking the derivative of $U_I$ and setting it to zero gives us the following differential equation for $\delta(\varepsilon)$:

\[
\dfrac{d \delta}{d \varepsilon} + \delta(\varepsilon)\frac{\lambda^{'}(\kappa)}{\lambda(\kappa)} = \frac{1}{B}\frac{u^{'}(\kappa)}{\lambda(\kappa)}
\]

Under the boundary condition that $\delta(\kappa)=0$. This represents a well defined first order differential equation, that has an straightforward unique solution determined by the boundary condition. Solving for $\delta(\varepsilon)$, we have:

\[
\delta(\varepsilon) = \dfrac{u^{'}(\kappa)}{B\lambda^{'}(\kappa)}\left(1 - e^{\frac{\lambda^{'}(\kappa)}{\lambda(\kappa)}(\kappa-\varepsilon)}\right)
\]

Combining all these results together  us the following lemma.

\begin{lemma}
 \label{lemma_m1}
 A Perfect Bayesian Equilibrium for the game when $|I|=1$ and $|C|=1$ is a pair of strategies and beliefs such that the optimal strategies are
 \begin{enumerate}
  \item For the incumbent
  \[
  \varepsilon^{*}(\eta) \ = \ \twopdo{\eta + \kappa}{\eta \leq \overline{\eta}}{A}
  \]
  \item For the challenger
  \[
  \varphi^*(\varepsilon) \ = \ \threepdo{0}{\varepsilon < \kappa}{\delta(\varepsilon)}{\varepsilon \in [\kappa, \overline{\eta}]}{\delta(A)}
  \]
  \item And for the voter, retain the incumbent if $\dd=0$ and $\psi=G$ and dismiss otherwise.
 \end{enumerate}
 With 
 \[
 \delta(\varepsilon) = \dfrac{u^{'}(\kappa)}{B\lambda^{'}(\kappa)}\left(1 - e^{\frac{\lambda^{'}(\kappa)}{\lambda(\kappa)}(\kappa-\varepsilon)}\right)
 \]
 and
 \[
 \kappa = \lambda^{-1}\left(\dfrac{c}{B}\right)
 \]
 And the optimal beliefs updated by Bayes rule whenever possible, with the out-of-the path beliefs restricted by the Universal Divinity Criteria.
\end{lemma}

\begin{proof}

Let $\varepsilon^*$ and $\varphi^*$ as defined in the Lemma \ref{lemma_m1}. To show the existence and the characteristics of the separating equilibrium, following \citet{Malaith1987} we need to prove the following Claim:

\begin{claimnn}\label{claim1}
The expenditure in weakly increasing in the type $t_i \in T$
\end{claimnn}

\begin{proof}[Proof of Claim \ref{claim1}]
First, notice that the type space $T$ is a compact set, and thus a complete lattice. Then:

\[
\dfrac{\partial^2 U}{\partial \varepsilon \partial \eta} = -u^{''} + B(\varphi(\varepsilon)\lambda^{''} + \varphi^{'}(\varepsilon)\lambda^{'}) 
\]

And so, as $\lambda$ is weakly increasing, $\lambda^{''}>0$ and $u$ is concave, $U$ satisfies the increasing differences properties in $\varepsilon$ and $t$. As $T$ is a lattice, we have that $\varepsilon^{*}(\eta)$, which is a maximum for $U_I(t_i, \varepsilon(t_i) )$, is also weakly increasing in $\eta$.\footnote{See \citet{MilgromShannon1994}, theorem 6.}

\end{proof}

As I derived the optimal strategies for the incumbent and challenger's strategies, it remains to show that no type $t_i \in T$ have incentives to deviate. Define the profit that a type $t_i$ gains when she deviates from $\varepsilon(t_i)$ to $\tilde{\varepsilon}$ as:

\[
p(t_i, \varepsilon(t_i), \varphi, \tilde{\varepsilon}) \ = \ u (\varepsilon(t_i) - \eta) + B \left( \varphi(\tilde{\varepsilon})\lambda(\varepsilon(t_i)-\eta) - \varphi(\varepsilon(t_i))\lambda(\varepsilon(t_i)-\eta) \right)
\]

Now consider the following Claim:

\begin{claimnn}\label{claim2}
No type has incentive to deviate from the equilibrium enunciated in Lemma \ref{lemma_m1}.
\end{claimnn}

\begin{proof}[Proof of the Claim \ref{claim2}]

Let $t_i = \eta$. We then have that

\[
\dfrac{\partial p}{\partial \eta} \ = \ \left(\dfrac{\partial \varepsilon}{\partial \eta} - 1 \right) (u^{'}(\varepsilon-\eta) - B\varphi(\varepsilon)\lambda^{'}(\varepsilon-\eta))
 - B \varphi(\tilde{\varepsilon})\lambda^{'}(\varepsilon-\eta)) - \dfrac{\partial \varepsilon}{\partial \eta} B \varphi^{'}(\varepsilon)\lambda(\tilde{\varepsilon}-\eta) - u^{'}(\varepsilon-\eta)
\]

Applying the fact that $\dfrac{\partial}{\partial \varepsilon}\left[ \varphi(\varepsilon) \lambda(\varepsilon - \eta) \right] = \dfrac{u^{'}(\varepsilon(\tilde{\eta})-\eta)}{B}$:

\[
\dfrac{\partial p}{\partial \eta} \ = \ -u^{'}(\varepsilon-\eta) + B(\varphi(\varepsilon)\lambda^{'}(\varepsilon-\eta) - \varphi(\tilde{\varepsilon}) \lambda(\tilde{\varepsilon} - \eta))
\]

And as long as the challenger plays $\varphi^*$, that is increasing on $\varepsilon$, the $ \dfrac{\partial p}{\partial \eta} < 0 $. Thus, $t_i$ has no incentives to deviate from $\varepsilon(\eta)$.

\end{proof}

For the voter, the optimal $\bups$ is $\bups=(1,0)$ when $\psi=G$ and $\dd=0$ and $\bups=(0,1)$ otherwise.

The beliefs are derived using Bayes rule:

\[
\mu(\eta | \varepsilon) = \dfrac{f(\eta)}{\int_\kappa^\varepsilon f(\eta) d\eta}
\]

As no discontinuity is present inside the interval $[\kappa, \overline{\eta}]$, the set of players who deviate under these strategies is a measure zero set. So, $(\varepsilon^*, \ \varphi^*, \ \bups^*, \ \mu^* )$, as defined in the Lemma \ref{lemma_m1} are a Perfect Bayesian Equilibrium for the game.

\end{proof}

## Single-member District with Many Challengers

Let $|I|=1$ and $|C|=L$. Solving backwards, the optimal strategy for the voters is to retain the incumbent if the public goods were successfully provided and no corruption was successfully uncovered. If the voters dismiss the incumbent, they randomly select a challenger among all challengers.

One might think that an optimal scheme would be to reward the challengers that uncovered the corruption scandal. The problem with this consideration is twofold: first, once the corruption has been detected, all challengers could use these facts in their campaigns, making the corruption exposure a non-excludable public good. Second, even if we consider that challengers inherit the votes from the exposed incumbent, the collective action remains, as every challenger will have incentives for negative campaign, but if all do, the voter will keep choosing by chance. The equilibrium changes slightly but qualitatively, it is still worse than just one challenger.

Challengers observe $\psi$ and $\varepsilon$, and have to simultaneously compute the negative campaign investment $\varphi_i$. As challengers have the same utility function, we solve for a symmetric equilibrium.\footnote{As the payoffs and costs are symmetric here, the symmetric Nash equilibrium is the adequate choice for this problem. If we had heterogeneity on challengers, induced by ideological preferences of the electorate, or other characteristics, the symmetry assumption would not be well suited.}

Let $\lambda(\varepsilon-\eta) = \lambda_\varepsilon$. If we have $k$ players investing in negative campaign against the incument politician, the probability that the incumbent loses the election is:

\[
\Prob(\text{success} | \text{k exp.})  \ = \ 1 - (1-\lambda_\varepsilon)^k
\]

Let $\varphi$ the probability that all players but $i$, invest in negative campaign. Then, the chance of having $k$ players investing in exposing the corrupt incumbent among the $L-1$ players is:

\[
\Prob( \text{k exp.}) \ = \ {L \choose k} \varphi^k (1-\varphi)^{L-1-k}
\]

And the composite probability of losing an election is given by:

\[
f(\varphi, x) = \sum_{k=0}^{L-1} {L \choose k} \varphi^k (1-\varphi)^{L-1-k} \cdot (1 - (1-\lambda_\varepsilon)^{k+x})
\]

And when $i$ invests in negative campaign, we let $x=1$ in $f$. This allows us to compute the expected utility for the challenger $i$ as:

\[
\varphi_i \left[\frac{B}{L}\int_0^A f(\varphi, 1) \mu(\eta|\varepsilon) d\mu \right] + (1-\varphi_i)\left[\frac{B}{L} \int_0^A f(\varphi, 0) \mu(\eta|\varepsilon) d\mu \right] - \varphi_i c
\]

And some algebra yields to:

\[
\frac{B}{L}\int_0^A f(\varphi, 0) \mu(\eta|\varepsilon) d\mu + \varphi_i  \left[\frac{B}{L} \int_0^A \left( f(\varphi, 1)-f(\varphi, 0) \right) \mu(\eta|\varepsilon) d\mu - c \right]
\]

The challenger $i$ exposes the incumbent whenever the term in the right-hand side is positive. Solving for $f$, taking the difference and applying to the problem gives us:

\[
\frac{B}{L} \int_0^A \left( (1-\varphi\lambda_\varepsilon)^{L-1}\lambda_\varepsilon \right) \mu(\eta|\varepsilon) d\mu - c
\]

And the sequential rationality imposes us that $\varphi_i \geq 0$ when this quantity is positive, and $0$ otherwise. The upper bound in the equilibrium is:

\[
\dfrac{\int_{\bar{\eta}}^A \left( (1-\varphi\lambda_A)^{L-1}\lambda_A \right) f(\eta) d\eta}{\int_{\bar{\eta}}^A f(\eta) d\eta} = L\dfrac{c}{B}
\]

And yields to:

\[
\int_0^{\bar{\eta}} \left((1-\varphi\lambda_\varepsilon)^{L-1}\lambda_\varepsilon - L\dfrac{c}{B} \right) \mu(\eta|\varepsilon) d\mu = 0
\]

Let us denote as $G$ the following equation:

\[
G(\lambda, \varphi; L) = (1-\varphi\lambda_\varepsilon)^{L-1}\lambda_\varepsilon - Q(B, c) L
\]

Where $Q$ is defined as in the previous section: $Q(B,c) = \dfrac{c}{B}$. The optimal $\varepsilon$ can be derived from this equation. The problem is that we will have, for each value of $\varepsilon$ a corresponding optimal value of $\varphi$, and vice-versa. As the incumbent plays first, she can choose the max-min value for $\lambda$, which is the maximum value for $\varepsilon$ that still delivers the lowest $\varphi$ possible. For a numerical example, if we assume $Q=0.15$ and $L$ ranging from $1$ to $5$, we can have a contour for the possible values for $\lambda$ in Figure \ref{fig_m2_1}.

\begin{figure}[htb!]
 \centering 
 \includegraphics{1figs/figm21.pdf}
 \caption{Contour plot for $G=0$ when $Q=0.15$ and $L$ ranges from $1$ to $5$}
 \label{fig_m2_1}
\end{figure}

And we have the following optimal incumbent expenditure strategy:

\[
\varepsilon^*(\eta) \ = \ \eta + \lambda^{-1}\left( Q(B, c)\cdot L \right) \ = \ \eta + \kappa(L)
\]

Again, the incumbent strategy is comprised of the actual value required to provide the service plus a kickback. However, now the kickback is augmented by the number of challengers contesting the seat.

Again, this expenditure has to be incentive compatible in order to the incumbent prefer it. Thus, it has to satisfy the following differential equation in $\delta(\varepsilon)$:

\[
u^{'}(\varepsilon-\eta) - BL(1-\delta(\varepsilon)\lambda(\varepsilon-\eta))^{L-1} (\delta^{'}(\varepsilon)\lambda(\varepsilon-\eta)-\delta(\varepsilon)\lambda(\varepsilon-\eta)^{'}) = 0
\]

But making $G=0$ yield to $(1-\delta(\varepsilon)\lambda(\varepsilon-\eta))^{L-1} = L\frac{Q(B,c)}{\lambda(\varepsilon-\eta)}$. Applying these results on the previous equations leads to the following differential equation:

\[
\delta^{'}(\varepsilon)-\delta(\varepsilon)\dfrac{\lambda^{'}(\kappa(L))}{\lambda(\kappa(L))} \ = \ \dfrac{u^{'}(\kappa(L))}{BQ(B, c)L^2}
\]

With the boundary condition $\delta(\kappa(L))=0$. Thus:

\[
\delta(\varepsilon) \ = \ \dfrac{u^{'}(\kappa(L))}{BQ(B, c)L^2} \left( 1- e^{\frac{\lambda^{'}(\kappa(L))}{\lambda(\kappa(L))}(\kappa(L)-\varepsilon)}\right)
\]

\begin{lemma}
 \label{lemma_m2}
 A Perfect Bayesian Equilibrium for the game when $|I|=1$ and $|C|=L$ is a pair of strategies and beliefs such that the optimal strategies are
 \begin{enumerate}
  \item For the Incumbent
  \begin{equation*}
  \varepsilon^{*}(\eta) \ = \ \twopd{\eta + \kappa(L)}{\eta \leq \overline{\eta}}{A}{\eta > \overline{\eta}}
  \end{equation*}
  \item For the Challenger
  \[
  \varphi^*(\varepsilon) \quad = \quad \threepdo{0}{\varepsilon < \kappa(L)}{\delta(\varepsilon)}{\varepsilon \in [\kappa(L), \overline{\eta}]}{\delta(A)}
  \]
  \item And the voters retain the incumbent if $d=0$ and $\psi=G$.
 \end{enumerate}
 And the optimal beliefs are the same as in the previous Lemma.
\end{lemma}

\begin{proof}

The results in this Lemma follow analogously from the previous Lemma.

As we are defining a symmetric equilibrium, we have $\varphi_i = \varphi$, $\forall i \in \{1, \cdots, L\}$. Similarly to the previous Lemma, suppose we have an optimal deviation:

\[
\dfrac{\partial}{\partial \tilde{\eta}} = \dfrac{\partial \varepsilon}{\partial \tilde{\eta}} \left( u^{'}(\varepsilon(\tilde{\eta})-\eta) - B L (1-\varphi(\varepsilon)\lambda(\varepsilon(\tilde{\eta})-\eta))^{L-1} \dfrac{\partial}{\partial \varepsilon}(\varphi(\varepsilon)\lambda(\varepsilon(\tilde{\eta})-\eta)) \right) = 0
\]

And as $ \dfrac{\partial \varepsilon}{\partial \tilde{\eta}}>0$, we should have that

\[
\dfrac{\partial}{\partial \varepsilon} (\varphi (\varepsilon) \lambda (\varepsilon (\eta)-\eta)) = \dfrac{u^{'}(\varepsilon(\eta)-\eta)}{B L (1-\varphi(\varepsilon)\lambda(\varepsilon(\eta)-\eta))^{L-1}}
\]

And the profit from deviate is:

\[
p(t_i, \varepsilon(t_i), \tilde{\varepsilon}) \ = \ u(\varepsilon-\eta) + B\left[ (1-\varphi(\varepsilon)\lambda(\varepsilon-\eta))^L - (1-\varphi(\tilde{\varepsilon})\lambda(\tilde{\varepsilon}-\eta))^L\right]
\]

Taking the derivative yields to:

\[
\dfrac{\partial p}{\partial \eta} \ = \ u^{'}(\varepsilon-\eta) - LB(1-\varphi(\tilde{\varepsilon})\lambda(\tilde{\varepsilon}-\eta))^{L-1}\varphi(\tilde{\varepsilon})\lambda^{'}(\tilde{\varepsilon}) \ < \ 0
\]

Therefore, no type has incentives to deviate from the equilibrium. The beliefs are computed with a similar method from Lemma \ref{lemma_m1}.

\end{proof}

And as the key result for the paper is that more challengers represent a larger larger expenditure proposals:

\begin{claim}
 \label{claim_veps}
 Let $\eta$ fixed. $\varepsilon^*(\eta; \ L)$ is increasing on $L$
\end{claim}

\begin{proof}[Proof of Claim \ref{claim_veps}]

Taking the derivative of $\varepsilon^*(\eta; L)$ on $L$ yields to:

\begin{eqnarray*}
\dfrac{\partial \varepsilon^*(\eta; L)}{\partial L} & = & \dfrac{\partial}{\partial L} \left[ \eta + \lambda^{-1} (Q(B,c) \cdot L) \right] \\
& = & \frac{Q(B,c)}{\lambda^{'}(\kappa(L))}
\end{eqnarray*}

And as $\lambda^{'}>0$ and $Q>0$, then $\dfrac{\partial \varepsilon^*(\eta, L)}{\partial L}>0$.

\end{proof}

## Multi-member Districts with Many Challengers

Consider now an election with $K$ incumbents and $L$ challengers. Solving backwards, the voters reelect a given incumbent $i$ if the public goods were successfully provided, and no corruption was exposed.

Challengers then decide whom to expose with negative campaign. They can choose to expose each one of the $K$ incumbents. In a logic similar to the previous Lemmas, we have that the optimal investment is:

\[
\varepsilon^{*}(\eta) \ = \ \eta + \lambda^{-1} ((L+K)\dfrac{c}{B}) \ = \ \eta + \kappa(K+L)
\]

And $\delta(.)$ defined in the similar mode as before.

\begin{lemma}
 \label{lemma_m4}
 A Perfect Bayesian Equilibrium for the game when $|I|=K$ and $|C|=L$ will be such as the strategies for the players are:
 \begin{enumerate}
  \item For each of the incumbents:
  \[
  \varepsilon^{*}(\eta) \ = \ \twopdo{\eta + \kappa(L+K)}{\eta \leq \overline{\eta}}{A}
  \]
  \item For each of the challengers:
  \[
  \varphi^*(\varepsilon) \ = \ \threepdo{0}{\varepsilon < \kappa(L+K)}{\delta(\varepsilon)}{\varepsilon \in [\kappa(L+K), \overline{\eta}]}{\delta(A)}
  \]
  \item For the voters, retain the incumbents that $d=0$ when $\psi=G$, dismissing otherwise.
 \end{enumerate}
 And the optimal beliefs are uptadated by Bayes rule, whenever possible.
\end{lemma}

The proof is similar to the previous proofs, and is available upon request.

# Model Extensions

## Run-offs

Run-off works in this model as a two-stage election, with the first election a contest with one incumbent and many challengers, and then a contest with two politicians. The behavioral rule assumed for the voters imply here that voters will send the incumbent and challenger to the run-off if no corruption was exposed in the first round. In equilibrium, challengers take this into consideration, not investing any resources in negative campaign in the first round. However, the run-off emulates a single-member district election with the incumbent and a challenger. This system is the most effective in coping corruption, accoding to Lemma \ref{lemma_m1}.

## Parties

Strong parties can change the game substantially by coordinate the investigations. They can optimally decide to invest the resources aiming at maximizing the changes of exposure. The logic with district size and competition reverts to parties, instead of candidates. Few strong parties can successfully coordinate the exposures, lowering the equilibrium corruption.

# Empirical Analysis

## Data Sources

I use data from Brazil for the 2004 municipal election, coupled with the corruption investigation performed by a Federal Random Audit program. The corruption investigations are performed on federal transfers, but these compose the majority of the municipal budgets in Brazil. Table \ref{tabsourc} provide a list of data sources, that can easily searched online.

The summary statistics follow below:

```
require(foreign)
require(fBasics)
require(rdrobust)
require(AER)
require(tidyverse)
require(ggplot2)
require(broom)
require(gridExtra)

dat <- read.dta('EISC.dta')
dat$nCandCC_by10 <- dat$nCandCC/10
dat$nCandCC <- NULL
res <- data.frame(round(t(basicStats(dat)[c("Mean", "Stdev", "Median", "Minimum", "Maximum", "nobs"),]), digits=3))
names(res) <- c("Mean", "SD", "Med", "Min", "Max", "N")
res

stdzy <- function(x) {
  z <- (x-mean(x, na.rm = T))/sd(x, na.rm = T)
  return(z)
}

# Extract coeffs
extr_coef_rdr <- function(obj, name = 'RD Robust') {
  return(data.frame(term = name, estimate = round(obj$coef[3], digits=3), std.error = round(obj$se[3], digits=3) ))
}

extr_coef_ttest <- function(obj, name = 'T Test') {
  cfs = obj$estimate[2]-obj$estimate[1]
  stde = abs((cfs)*(1/obj$statistic))
  res <- data.frame(term = name, 
                    estimate = round(cfs, digits=3), 
                    std.error = round(stde, digits=3))
  row.names(res) <- NULL
  return(res)
}

extr_coef_ivreg <- function(obj, name = 'IV Regression') {
  return(data.frame(term = name, 
                    estimate = round(obj$coefficients[2,1], digits=3),
                    std.error = round(obj$coefficients[2,2], digits=3)))
}
```

And the data sources are:

\usepackage{lscape}
\begin{landscape}
\begin{table}[htp!]
\centering
\caption{Data sources used in empirical section}
\label{tabsourc}
\scalebox{0.8}{
\begin{tabular}{llcl}
\hline
\multicolumn{1}{c}{\textbf{Variable}} & \multicolumn{1}{c}{\textbf{Description}} & \textbf{Source} & \multicolumn{1}{c}{\textbf{URLs}} \\ \hline
ibge\_code & \begin{tabular}[c]{@{}l@{}}6-digit Brazilian Geography and Statistics \\ Institute municipal identifier\end{tabular} & IBGE & https://cidades.ibge.gov.br \\
year & Year &  &  \\
lottery & \begin{tabular}[c]{@{}l@{}}Number draw from CGU \\ (Brazilian General Comptroller)\end{tabular} & CGU & http://www.cgu.gov.br \\
broad & \begin{tabular}[c]{@{}l@{}}Brollo et al (2013) broad \\ detected irregularities\end{tabular} & \begin{tabular}[c]{@{}c@{}}CGU\\ + processing\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.aeaweb.org/articles?id=10.1257/aer.103.5.1759\\ http://www.cgu.gov.br\end{tabular} \\
narrow & \begin{tabular}[c]{@{}l@{}}Brollo et al (2013) corruption \\ detected\end{tabular} & \begin{tabular}[c]{@{}c@{}}CGU\\ + processing\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.aeaweb.org/articles?id=10.1257/aer.103.5.1759\\ http://www.cgu.gov.br\end{tabular} \\
ind\_corrup\_n & \begin{tabular}[c]{@{}l@{}}Avelino et al (2016) fraction of detected \\ irregularities\end{tabular} & \begin{tabular}[c]{@{}c@{}}CGU\\ + processing\end{tabular} & https://doi.org/10.7910/DVN/MJWO0X \\
ind\_corrup\_m & \begin{tabular}[c]{@{}l@{}}Avelino et al (2016) monetary amount \\ of detected irregularities\end{tabular} & \begin{tabular}[c]{@{}c@{}}CGU\\ + processing\end{tabular} & https://doi.org/10.7910/DVN/MJWO0X \\
nCandMayor & \begin{tabular}[c]{@{}l@{}}Number of candidates running for \\ 2004 mayoral election\end{tabular} & TSE & http://www.cepesp.io \\
runningCC & \begin{tabular}[c]{@{}l@{}}Distance from 2003 TSE council \\ size decision and 2003 IBGE population\end{tabular} & \begin{tabular}[c]{@{}c@{}}TSE + IBGE\\ + processing\end{tabular} & http://www.cepesp.io \\
fakeRunningCC & \begin{tabular}[c]{@{}l@{}}Placebo distance from 2003 TSE \\ council size decision and 2003 IBGE population\end{tabular} & \begin{tabular}[c]{@{}c@{}}TSE + IBGE\\ + processing\end{tabular} & http://www.cepesp.io \\
pop & \begin{tabular}[c]{@{}l@{}}Population projection for the \\ research period\end{tabular} & IBGE & https://cidades.ibge.gov.br \\
validVotes\_by100k & \begin{tabular}[c]{@{}l@{}}Number of voters by municipality \\ (divided by 100 thousands)\end{tabular} & TSE & http://www.tse.jus.br \\
distValidVotes200k & \begin{tabular}[c]{@{}l@{}}Distance from the 200 thousands \\ run-off cutoff\end{tabular} & TSE + processing & http://www.tse.jus.br \\
distValidVotes150k & \begin{tabular}[c]{@{}l@{}}Placebo distance from the 200 \\ thousands run-off cutoff (placebo as 150k)\end{tabular} & TSE + processing & http://www.tse.jus.br \\
distValidVotes250k & \begin{tabular}[c]{@{}l@{}}Placebo distance from the 200 \\ thousands run-off cutoff (placebo as 250k)\end{tabular} & TSE + processing & http://www.tse.jus.br \\
year\_2005 & Dummy for year = 2005 &  &  \\
year\_2006 & Dummy for year = 2006 &  &  \\
year\_2007 & Dummy for year = 2007 &  &  \\
year\_2008 & Dummy for year = 2008 &  &  \\
nCandCC\_by10 & \begin{tabular}[c]{@{}l@{}}Number of candidates running \\ for city council (divided by 10)\end{tabular} & TSE & http://www.tse.jus.br \\ \hline
\end{tabular}
}
\end{table}
\end{landscape}

## Estimated Models

I run three models. First, a RDD for the city council size. Second, I tried a RDD for the run-off thresholds, but the number of cases on the right hand side is too small, and the RDD is unidentified. To avoid this problem, I run a t-test comparing the cases on the left with the cases on the right, restricting the sample for the municipalities around 50 thousands away from the cutoff (for identification purposes).

For \citet{brollo2013political} broad irregularities:

```
trs = .5 # Chosen to make the computations feasible
aux <- rdrobust(dat$broad, dat$runningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008', 'lottery')])

coefs <- extr_coef_rdr(aux, name = 'Council Size (RDD)')

aux <- t.test(dat$broad[abs(dat$distValidVotes200k)<trs]~as.numeric(dat$distValidVotes200k[abs(dat$distValidVotes200k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'Run-off (T-test)'))

aux <- summary(ivreg(broad~stdzy(nCandMayor)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Mayor (IVreg)'))

aux <- summary(ivreg(broad~stdzy(nCandCC_by10)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Council (IVreg)'))

coefs
coefs$model = 'Broad inefficiencies'
totalModel <- coefs
```

For \citet{brollo2013political} narrow (corruption) irregularities:

```
aux <- rdrobust(dat$narrow, dat$runningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008', 'lottery')])

coefs <- extr_coef_rdr(aux, name = 'Council Size (RDD)')

aux <- t.test(dat$narrow[abs(dat$distValidVotes200k)<trs]~as.numeric(dat$distValidVotes200k[abs(dat$distValidVotes200k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'Run-off (T-test)'))

aux <- summary(ivreg(narrow~stdzy(nCandMayor)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Mayor (IVreg)'))

aux <- summary(ivreg(narrow~stdzy(nCandCC_by10)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Council (IVreg)'))
coefs

coefs$model = 'Narrow corruption'
totalModel <- rbind(coefs, totalModel)
```

```
p <- ggplot(totalModel, aes(term, estimate))+
  geom_point()+
  geom_pointrange(aes(ymin = estimate - 1.64*std.error, 
                      ymax = estimate + 1.64*std.error), size = 0.5)+
  geom_pointrange(aes(ymin = estimate - 1.96*std.error, 
                      ymax = estimate + 1.96*std.error), 
                  size = 0.35, linetype = 2)+
  geom_text(aes(label = round(estimate, digits=2)), nudge_x = 0.25) + 
  scale_x_discrete("") +
  theme_bw() +
  geom_hline(yintercept = 0, linetype = 3) +
  coord_flip(ylim = c(-1, 1)) + ylab(NULL)+
  theme(text = element_text(size=12))+
  facet_grid(~model)
```

For \citet{avelino2016replication} incidence of corruption in health care transfers:

```
aux <- rdrobust(dat$ind_corrup_n, dat$runningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008', 'lottery')])

coefs <- extr_coef_rdr(aux, name = 'Council Size (RDD)')

aux <- t.test(dat$ind_corrup_n[abs(dat$distValidVotes200k)<trs]~as.numeric(dat$distValidVotes200k[abs(dat$distValidVotes200k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'Run-off (T-test)'))

aux <- summary(ivreg(ind_corrup_n~stdzy(nCandMayor)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Mayor (IVreg)'))

aux <- summary(ivreg(ind_corrup_n~stdzy(nCandCC_by10)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Council (IVreg)'))

coefs
coefs$model = 'Pct. corruption health'
totalModel <- coefs
```

For \citet{avelino2016replication} amount of corruption in health care transfers (divided by the total transfers):

```
aux <- rdrobust(dat$ind_corrup_m, dat$runningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008', 'lottery')])

coefs <- extr_coef_rdr(aux, name = 'Council Size (RDD)')

aux <- t.test(dat$ind_corrup_m[abs(dat$distValidVotes200k)<trs]~as.numeric(dat$distValidVotes200k[abs(dat$distValidVotes200k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'Run-off (T-test)'))

aux <- summary(ivreg(ind_corrup_m~stdzy(nCandMayor)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Mayor (IVreg)'))

aux <- summary(ivreg(ind_corrup_m~stdzy(nCandCC_by10)+pop+factor(year)|runningCC+pop+factor(year), data=dat), vcov = sandwich, df = Inf, diagnostics = TRUE)

coefs <- rbind(coefs, extr_coef_ivreg(aux, name = 'Cand. Council (IVreg)'))

coefs
coefs$model = 'Tot. Amt. corrup. health'
totalModel <- rbind(coefs, totalModel)
```

And the coefficient plots:

```
p2 <- ggplot(totalModel, aes(term, estimate))+
  geom_point()+
  geom_pointrange(aes(ymin = estimate - 1.64*std.error, 
                      ymax = estimate + 1.64*std.error), size = 0.5)+
  geom_pointrange(aes(ymin = estimate - 1.96*std.error, 
                      ymax = estimate + 1.96*std.error), 
                  size = 0.35, linetype = 2)+
  geom_text(aes(label = round(estimate, digits=2)), nudge_x = 0.25) + 
  scale_x_discrete("") +
  theme_bw() +
  geom_hline(yintercept = 0, linetype = 3) +
  coord_flip(ylim = c(-1, 1)) + ylab(NULL)+
  theme(text = element_text(size=12))+
  facet_grid(~model)

grid.arrange(p, p2, ncol = 1)
```

## Robustness Checks

For the RDD estimates I run the same models, now with a fake cutoff. The results follow below:

```
aux <- rdrobust(dat$broad, dat$fakeRunningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008')])

coefs <- extr_coef_rdr(aux, name = 'broad')

aux <- rdrobust(dat$narrow, dat$fakeRunningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008')])

coefs <- rbind(coefs, extr_coef_rdr(aux, name = 'narrow'))

aux <- rdrobust(dat$ind_corrup_n, dat$fakeRunningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008')])

coefs <- rbind(coefs, extr_coef_rdr(aux, name = 'freqCorrupHealth'))

aux <- rdrobust(dat$ind_corrup_m, dat$fakeRunningCC, covs = dat[,c('pop','year_2006','year_2007','year_2008')])

coefs <- rbind(coefs, extr_coef_rdr(aux, name = 'amtCorrupHealth'))

p <- ggplot(coefs, aes(term, estimate))+
  geom_point()+
  geom_pointrange(aes(ymin = estimate - 1.64*std.error, 
                      ymax = estimate + 1.64*std.error), size = 0.5)+
  geom_pointrange(aes(ymin = estimate - 1.96*std.error, 
                      ymax = estimate + 1.96*std.error), 
                  size = 0.35, linetype = 2)+
  geom_text(aes(label = round(estimate, digits=2)), nudge_x = 0.25) + 
  scale_x_discrete("") +
  theme_bw() +
  geom_hline(yintercept = 0, linetype = 3) +
  coord_flip(ylim = c(-1, 1)) + ylab(NULL)+
  theme(text = element_text(size=12))
p
```

And as we can see, most of the placebo RDD tests failed, as expected. The only significant result is the amount of health care corruption, and in the wrong direction. For the run-off cutoffs, we did not use RDD because the number of cases in each side is small. We run a robustness with a fake cutoff for 150 thousand (50 thousand voters below the actual threshold). The results follow below:

```
aux <- t.test(dat$broad[abs(dat$distValidVotes150k)<trs]~as.numeric(dat$distValidVotes150k[abs(dat$distValidVotes150k)<trs]>0))

coefs <- extr_coef_ttest(aux, 'broad (below)')

aux <- t.test(dat$narrow[abs(dat$distValidVotes150k)<trs]~as.numeric(dat$distValidVotes150k[abs(dat$distValidVotes150k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'narrow (below)'))

aux <- t.test(dat$ind_corrup_n[abs(dat$distValidVotes150k)<trs]~as.numeric(dat$distValidVotes150k[abs(dat$distValidVotes150k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'freqCorrupHealth (below)'))

aux <- t.test(dat$ind_corrup_m[abs(dat$distValidVotes150k)<trs]~as.numeric(dat$distValidVotes150k[abs(dat$distValidVotes150k)<trs]>0))

coefs <- rbind(coefs, extr_coef_ttest(aux, 'amtCorrupHealth (below)'))

p <- ggplot(coefs, aes(term, estimate))+
  geom_point()+
  geom_pointrange(aes(ymin = estimate - 1.64*std.error, 
                      ymax = estimate + 1.64*std.error), size = 0.5)+
  geom_pointrange(aes(ymin = estimate - 1.96*std.error, 
                      ymax = estimate + 1.96*std.error), 
                  size = 0.35, linetype = 2)+
  geom_text(aes(label = round(estimate, digits=2)), nudge_x = 0.25) + 
  scale_x_discrete("") +
  theme_bw() +
  geom_hline(yintercept = 0, linetype = 3) +
  coord_flip(ylim = c(-1, 1)) + ylab(NULL)+
  theme(text = element_text(size=12))
p
```

And the placebo failed, as expected, for all models. We choose a window of 50 thousand voters, as restricting more would make the estimation infeasible.

\clearpage