### 2021-08-24 article/article.Rmd
intro advancing

---
title:  |
        | Electoral Systems, Competition, and Incentives for Corruption
author: 
  - Umberto Mignozzetti[^umberto]
date: \today
thanks: "I thank Livio DiLonardo, Sandy Gordon, Saad Gulzar, Cathy Hafer, Carlo Holz, Dimitri Landa, Adam Przeworski, Howard Rosenthal, Renard Sexton, Denis Stukal, and Scott Tyson for their comments. I also thank participants at APSA, MPSA, and SPSA. All materials required to replicate this study are available at <https://github.com/umbertomig/electoral-systems-corruption>."
abstract: "What are the effects of electoral systems and competition on political corruption? In this paper, I propose a formal model of political corruption where incumbents decide how much kickbacks to add in the XXXX, and challengers may run anti-corruption campaigns against the incumbents. To understand how institutions and competition affect the results, I study two benchmarks: in the first, called the non-excludable anti-corruption campaign, successful exposure of an incumbent opens her seats to all other competitors. In the second benchmark, the excludable anti-corruption campaign, the politician successfully exposing the corruption signals competence, and voters tend to prefer her vis-a-vis the other challengers. When anti-corruption campaigns are non-excludable, an increase in the number of challengers or incumbents monotonically raises corruption. However, when anti-corruption campaigns are excludable, the number of challengers is endogenously limited. The excludability makes the free-riding in corruption exposure disappear. When I increase the number of incumbents, corruption rises less than in the non-excludable benchmark. These predictions hold empirically using quasi-experimental data from Brazilian and Italian municipalities. This study has implications for the design of electoral institutions and entry barriers in democratic elections."
abstractspacing: double
keywords: Political Corruption; District Magnitude; Single-member districts; Multi-member districts; Welfare; Brazil
#jelcodes: HXXXX
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# Knitr options
knitr::opts_chunk$set(fig.pos = "H") # holds figure position
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# XXXX
```

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>.

\newpage

# Introduction
\label{sec:intro}

Corruption is one of the most prevalent and harmful diseases of developing countries. Corruption lowers public goods provision \citep{banerjee201227, beekman2014corruption}, decreases foreign investments \citep{habib2002corruption}, stunt economic growth \citep{mauro1995corruption, grundler2019corruption}, and affects human capital investments \citep{ferraz2012corrupting}. Corruption also have detrimental effects on trust \citep{you2018trust}, attitudes toward political parties \citep{winters2015political}, and fuels the distrust in the political system \citep{mignozzetti2019brazil}. The International Transparency estimates that corruption is responsible for $1.23 trillion US dollars every year, a sum that could lift 1.4 billion people out of the extreme poverty for consecutive six years \citep{fleming2019corruption}.

One crucial distinction in corruption activity is the actor that is perpetrating corruption. While bureaucratic corruption may be more frequent, it is also easier to sanction and to punish. Political corruption, on the other hand, involves higher rents than bribes to services. They are usually associated with large government contracts and procurement \citep{broms2019political}, have a more disruptive effects on the trust in political institutions \citep{sole2018trust}, divert the bureaucracy toward rent seeking \citep{brierley2020unprincipled}. For instance, one of the major corruption scandals in the world, the Car Wash scandal in Brazil, involved more than 50 politicians, several high level bureaucrats, and about five billion US dollars in kickbacks \citep{watts2017operation}. Political corruption is intricate also because the electoral sanctioning often fails \citep{klavsnja2017uninformed, canete2020voting, janezic2020eliciting}. However, as elections are the primary mechanism to sanction corrupt politicians, are there electoral system features that could improve the sanctioning and lower corruption?

In this paper, I investigate the interplay between the number of incumbents in a polity, the number of competitors challenging those incumbents, and the incentives incumbents have for engaging in corruption. I propose a model where incumbents know the optimal level of investment required for public goods provision, but challengers only see the amount that incumbents spend. Incumbents have privileged information because they are inside the government, and receive information from governmental agencies. When they propose an investment higher than the optimal amount, the public goods is provided, and they pocket the difference between the optimal amount and their proposed expenditure. Thus, in this paper, the difference between the optimal and the proposed investment is the definition of political corruption.

Challengers have incomplete information about the required investment in public goods provision, but they may run a negative campaigning against incumbents. Challengers have a prior expectation regarding the optimal investments, and naturally, they know that the higher the level of investment proposed, the more the chances that incumbents are extracting rents. They decide to run a negative campaign that aims to gather and expose information supporting that a given incumbent is corrupt. To run a negative campaign is costly, both in terms of gathering information to prove that the incumbent is dishonest, and also in terms of XXXX. The extent through which the negative campaign is successful depends on the rents extracted by incumbents: While voters know that challengers may accuse an honest incumbent of corruption, the amount of evidence shown to voters can tilt the perception that a given incumbent is honest, triggering his or her dismissal.

To understand how electoral systems and competition affects corruption, I investigate how changes in the number of incumbents (district magnitude) and the number of challengers (competition) affect the equilibrium level of corruption. I show that increase district magnitude increases the equilibrium level of political corruption. This because more incumbents promote coordination failures in the challengers' negative campaigning. For example, in an electoral system with one incumbent, this incumbent is the sole target of negative campaigning. In a system with two incumbents, one incumbent may be targeted with all negative campaign while the other may remain un-targeted. These types of coordination shortcomings make the incumbents to propose higher expenditures, increasing corruption.

The effects of competition are less clear, and depend on the how voters behave in face of a successful exposure of an incumbent. On the one hand, if voters transfer the political support from the corrupt incumbent to the challenger that ran the negative campaign, then political corruption decreases. This because the reward in denouncing a corrupt incumbent makes challengers more prone to run negative campaigns. With more negative campaigning in equilibrium, incumbents respond by proposing an expenditure closer to the optimal spending level. On the other hand, if voters, upon being convinced that the incumbent is corrupt, decide to randomly elect one challenger, then political corruption increases. This because randomly select the challenger that substitute the corrupt incumbent creates a free-riding incentive: every challenger prefers that other challengers pay the cost of negative campaign. The first mechanism we denote as XXXX while the second mechanism we denote as XXXX.

This model innovates in three fronts. First, it provides a consistent and comparable definition of corruption. Despite the evidence that corruption is detrimental for XXXX and XXXX, corruption is poorly defined. Corruption XXXX. Moreover, corruption XXXX.

Second, our paper builds up on the previous literature on electoral systems and corruption by adding a role for the challengers. In most of the papers, XXXX. My paper innovates by XXXX.

Third, my model abstracts away from other political concerns, focusing only on features of the electoral system, in a stylized model. Recent papers show that corruption correlates with ideology \citep{XXXX}, in-group behavior \citep{XXXX}, and XXXX. Despite the importance of these XXXX, removing these considerations is fundamental for understanding how electoral rules affect corruption. For example, by having ideology and district size in the same model, we would be uncertain whether which component (or even if the interaction between them) leads to corruption. Moreover, high polarization correlates with less punishment of corruption \citep{XXXX}, but it is unrealistic to prescribe that voters should depolarize as a way to lower corruption. The relation between polarization, XXXX, and XXXX are features of the political system that need to be studied in and on themselves. The strength of my model is to abstract away from XXXX, and considering a lean model of XXXX.

Finally, XXXX

# District Magnitude, Competition, and Corruption

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}